---
title: "NCSA"
author: "Guangya Wan"
date: "June 25, 2019"
output:
  html_document:
    df_print: paged
---
---
```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.align = "center")
```
# NCSA_WNV Report

## Report outline

I will use statistical model to explore the relationship between social_economical data with the West Nile Virus(approximated by MIR data in this case). I will try linear regression first here, testing and helpfully improving the model by some sort of transformation and selection. I will try other methods later and will update as soon as I get to those steps

## Linear regression

### Summary statistics and visulization

```{r}
data = read.csv('data.csv')
```

```{r}
data[1:10,] # first ten rows of data
```


```{r}
nrow(data) # number of rows of data
mean(data[['Annual_MIR']]) # mean of response
var(data[['Annual_MIR']]) # variance of response
min(data$Annual_MIR)
max(data$Annual_MIR)
hist(data$Annual_MIR,
      xlab   = "Severity",
     main   = "Histogram of MIR data",
     breaks = 100,
     col    = "dodgerblue",
     border = "darkorange") # Histogran of response
```

```{r}
library(faraway)
round(cor(data[-181]), 2)[1,1:10] # Correlation of data; as shown in the output, the correlation is large
```
```{r}
linear_model = lm(Annual_MIR ~ ., data = data)
# summary(linear_model)
```
```{r}
summary(linear_model)$r.squared ## R^2 of this model
summary(linear_model)$sigma ## Residual standard error
```
```{r}
vf = vif(linear_model)
```
```{r}
good_col = t(data.frame(vf[vf<=5]))
```


```{r}
new_data = data[colnames(good_col)]
new_data = cbind(new_data,data['Annual_MIR'])
```

```{r,eval=FALSE}
write.csv(subset(new_data,subset = mod_cd < 4 / length(mod_cd)),'selected_features_remove_outliers.csv')
```


```{r}
round(cor(new_data[-181]), 2)[1,1:10]
```

```{r}
new_linear_model = lm(Annual_MIR ~ ., data = new_data)
```
```{r}
summary(new_linear_model)$r.squared ## R^2 of this model
summary(new_linear_model)$sigma ## Residual standard error
```
```{r}
mod_cd = cooks.distance(new_linear_model)
sum(mod_cd > 4 / length(mod_cd))
linear_model_out = lm(Annual_MIR ~ ., data = data,subset = mod_cd < 4 / length(mod_cd))
```
```{r}
summary(linear_model_out)$r.squared ## R^2 of this model
summary(linear_model_out)$sigma ## Residual standard error
```


```{r}
library(leaps)
linear_sub = summary(regsubsets(Annual_MIR ~ ., data  = new_data,really.big = TRUE)) # selected features based on BIC
```
```{r}
best_adjr_col = which(linear_sub$which[which.max(linear_sub$adjr2),])
```

```{r}
adjr_data = cbind(new_data[colnames(t(data.frame(best_adjr_col)))[-1]],data['Annual_MIR'])
```
```{r}
adjr_model =  lm(Annual_MIR ~ ., data = adjr_data)
summary(adjr_model)
```


```{r}
plot(adjr_model)
```

```{r}
library(MASS)
library(faraway)
boxcox(adjr_model, plotit = TRUE, lambda = seq(-1.5, 1.5, by = 0.1))
```

```{r}
boxcox_bic = lm(((Annual_MIR ^ 0.15 - 1)  / 0.15) ~ ., data = adjr_data)
```

```{r}
plot(fitted(boxcox_bic), resid(boxcox_bic), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Data from Model 3")
abline(h = 0, col = "darkorange", lwd = 2)
qqnorm(resid(boxcox_bic), main = "Normal Q-Q Plot, box cox transformation", col = "darkgrey")
qqline(resid(boxcox_bic), col = "dodgerblue", lwd = 2)
```
```{r}
mod_cd = cooks.distance(boxcox_bic)
sum(mod_cd > 4 / length(mod_cd))
boxcox_bic = lm(((Annual_MIR ^ 0.2 - 1)  / 0.2) ~ ., data = adjr_data,subset = mod_cd < 4 / length(mod_cd) )
```

```{r}
plot(boxcox_bic)
```
```{r}
summary(boxcox_bic)
```

```{r}
set.seed(101) # Set Seed so that same sample can be reproduced in future also
# Now Selecting 75% of data as sample from total 'n' rows of the data  
sample <- sample.int(n = nrow(data), size = floor(.8*nrow(adjr_data)), replace = F)
train <- adjr_data[sample, ]
test  <- adjr_data[-sample, ]
```


```{r}
library(caret)
model <- train(
  (Annual_MIR ^ 0.2 - 1)  / 0.2 ~ ., 
  train,
  method = "lm",
  trControl = trainControl(
    method = "cv", 
    number = 5,
    verboseIter = TRUE,
  )
)
```
```{r}
summary(model)
```
```{r}
pred =predict(model,test)
pred[1:10]
test$Annual_MIR[1:10]
r_squared(test$Annual_MIR,pred)
```

